---
title: "Heart Disease UCI"
output: html_notebook
---

Lets load the required packages first

```{r}
library(readr) #to load dataset
library(tidyverse) #`tidyverse` contains packages like dplyr and ggplot which helps in exploration
library(ggthemes) #plot themes
library(corrplot) #for correlation plot
library(randomForest) #for predictive analysis
```

Now lets load the data and take a peak

```{r}
library(readr)
heart <- read_csv("heart.csv")
glimpse(heart)
```

Here lets take a moment to understand the variables :

<p><b>[01] age</b>       - Age in Years </p>
<p><b>[02] sex</b>       - 1: Male, 0: Female </p>
<p><b>[03] cp</b>        - Chest Pain (3 Levels -  1: Typical Angina, 2: Atypical Angina, 3: Non-Anginal Pain, 4: Asymptomatic) </p>
<p><b>[04] trestbps</b>  - Resting Blood Pressure (mm Hg; On Admission to the Hospital) </p>
<p><b>[05] chol</b>      - Cholestrol Level (mg/dl) </p>
<p><b>[06] fbs</b>       - Fasting Blood Sugar (> 120 mg/dl; 1: TRUE, 0: FALSE) </p>
<p><b>[07] restecg</b>   - Resting Electrocardiographic Measurement (0: Normal, 1: Having ST-T Wave Abnormality, 2: Showing Probable or Definite Left Ventricular Hypertrophy by Estes' Criteria) </p>
<p><b>[08] thalach</b>   - Maximum Heart Rate Achieved </p>
<p><b>[09] exang</b>     - Exercise Induced Angina (1: Yes, 0: No) </p>
<p><b>[10] oldpeak</b>   - ST Depression Induced by Exercise Relative to Rest('ST' denotes the positions on ECG Graph) </p>
<p><b>[11] slope</b>     - Slope of the Peak Exercise ST Segment (1: Upsloping, 2: Flat, 3: Downsloping) </p>
<p><b>[12] ca</b>        - Number of Major Vessels (0-3) </p>
<p><b>[13] thal</b>      - A Blood Disorder Called Thalassemia (3: Normal, 6: Fixed Defect, 7: Reversable Defect) </p>
<p><b>[14] target</b>    - Heart Disease (0: NO, 1: Yes) </p>

Lets change the column names to make it easier to read the data

```{r}
colnames(heart)[1:9] <- c("Age", "Sex", "Chest_Pain", "Resting_BP", "Cholestrol", "Fasting_Blood_Sugar",
                           "Resting_ECG", "Max_Heart_Rate", "Excercise_Induced_Angina")
colnames(heart)[12:14] <- c("Num_Major_Vessels", "Thalassemia", "Heart_Disease")
head(heart, n = 10)
```

Next lets see the correlation between these attributes

```{r}
heartcor <- cor(heart)
corrplot(heartcor, method = "pie", type = "lower")
```

<p>Here we find the following :</p>

<ol>
 <li>Chest pain and maximum heart rate has positive correlation to the presence of heart disease.</li>
 <li>While, excercise induced angima, oldpeak has negative correlation.</li>
</ol>

Let's do a bit more exploration and find out the percentage of people with heart disease in this dataset :

```{r}
round(table(heart$Heart_Disease)/nrow(heart), digits = 2)
```

So about 54% has heart disease and 46% does not have a heart disease in this dataset.

Now let's start with analysing :

first the data needs to be split

```{r}
sample.ind <- sample(2, nrow(heart), replace = T, prob = c(0.6,0.4))

heart.train <- heart[sample.ind==1,]
heart.test <- heart[sample.ind==2,]
```

Now let's check the heart disease probabilities in the now split dataset :

```{r}
round(table(heart.train$Heart_Disease)/nrow(heart.train), digits = 2)
round(table(heart.test$Heart_Disease)/nrow(heart.test), digits = 2)
```

Now let's begin building the classification random forest :

first the formula needs to be made -

```{r}
attributes <- names(heart)
attributes <- attributes[!attributes %in% c("Heart_Disease")]
attributes1 <- paste(attributes, collapse = "+")
formula <- as.formula(paste("Heart_Disease", attributes1, sep = " ~ "))
```

Now we can build the forest model, let's make 500 decision trees : 

```{r}
model.rf <- randomForest(formula, heart.train,
                              ntree=1000,
                              importance=T)
plot(model.rf)
legend("top", colnames(model.rf$err.rate),col=1:3,cex=0.8,fill=1:3)
```

```{r}
varImpPlot(model.rf,
           sort = T,
           main="Variable Importance",
           n.var=5)
```

```{r}
var.imp <- data.frame(importance(model.rf, type=2))
var.imp$Variables <- row.names(var.imp)
var.imp[order(var.imp$MeanDecreaseGini,decreasing = T),]
```

