---
title: "Heart Disease UCI"
output: html_notebook
---

Lets load the required packages first

```{r}
library(readr) #to load dataset
library(tidyverse) #`tidyverse` contains packages like dplyr and ggplot which helps in exploration
library(ggthemes) #plot themes
library(corrplot) #for correlation plot
```

Now lets load the data and take a peak

```{r}
library(readr)
heart <- read_csv("heart.csv")
glimpse(heart)
```

Here lets take a moment to understand the variables :

<p><b>[01] age</b>       - Age in Years </p>
<p><b>[02] sex</b>       - 1: Male, 0: Female </p>
<p><b>[03] cp</b>        - Chest Pain (3 Levels -  1: Typical Angina, 2: Atypical Angina, 3: Non-Anginal Pain, 4: Asymptomatic) </p>
<p><b>[04] trestbps</b>  - Resting Blood Pressure (mm Hg; On Admission to the Hospital) </p>
<p><b>[05] chol</b>      - Cholestrol Level (mg/dl) </p>
<p><b>[06] fbs</b>       - Fasting Blood Sugar (> 120 mg/dl; 1: TRUE, 0: FALSE) </p>
<p><b>[07] restecg</b>   - Resting Electrocardiographic Measurement (0: Normal, 1: Having ST-T Wave Abnormality, 2: Showing Probable or Definite Left Ventricular Hypertrophy by Estes' Criteria) </p>
<p><b>[08] thalach</b>   - Maximum Heart Rate Achieved </p>
<p><b>[09] exang</b>     - Exercise Induced Angina (1: Yes, 0: No) </p>
<p><b>[10] oldpeak</b>   - ST Depression Induced by Exercise Relative to Rest('ST' denotes the positions on ECG Graph) </p>
<p><b>[11] slope</b>     - Slope of the Peak Exercise ST Segment (1: Upsloping, 2: Flat, 3: Downsloping) </p>
<p><b>[12] ca</b>        - Number of Major Vessels (0-3) </p>
<p><b>[13] thal</b>      - A Blood Disorder Called Thalassemia (3: Normal, 6: Fixed Defect, 7: Reversable Defect) </p>
<p><b>[14] target</b>    - Heart Disease (0: NO, 1: Yes) </p>

Lets change the column names to make it easier to read the data

```{r}
colnames(heart)[1:9] <- c("Age", "Sex", "Chest_Pain", "Resting_BP", "Cholestrol", "Fasting_Blood_Sugar",
                           "Resting_ECG", "Max_Heart_Rate", "Excercise_Induced_Angina")
colnames(heart)[12:14] <- c("No._Major_Vessels", "Thalassemia", "Heart_Disease")
head(heart, n = 10)
```

Next lets see the correlation between these attributes

```{r}
heartcor <- cor(heart)
corrplot(heartcor, method = "pie", type = "lower")
```

<p>Here we find the following :</p>

<ol>
 <li>Chest pain and maximum heart rate has positive correlation to the presence of heart disease.</li>
 <li>While, excercise induced angima, oldpeak has negative correlation.</li>
</ol>